Formalizing the Kolmogorov extension Theorem in Lean

### 2023-09-10 Manuscript/proof_outline.tex
updates on the manuscript

\begin{theorem}[Kolmogorov extension]\label{T1}
  For all $t\in\iota$, let $\alpha_t$ be a separable, complete
  pseudo-extended-metric space and $\mathcal F_t$ the Borel
  $\sigma$-algebra generated by its topology. Let $(P_J)_{J\subseteq_f
    \iota}$ be a projective family of finite measures and $P$ be
  defined on $\mathcal A := \bigcup_{J \subseteq_f \iota} \mathcal
  F_J$ given by $P(A) = P_J(A)$ for $A\in\mathcal F_J$. Then, there is
  a unique extension of $P$ to $\sigma(\mathcal A)$.
\end{theorem}

@@

\noindent
In the proof of Kolmogorov's extension theorem, we use a closely
related definition of \leanline{Measure.ofAddContent}, based on
\leanline{ [mα : MeasurableSpace α] (hC : SetSemiring C) (hC_gen :
  MeasurableSpace.generateFrom C = mα)}, i.e.\ the semi-ring
\leanline{C} generates the $\sigma$-algebra, and which uses an
\leanline{m : AddContent} as well as \leanline{m_sigma_subadd} as
above.

\subsection{$\sigma$-additivity of set functions}
For the proof of Theorem~\ref{T1}, note that $P$ as given in the
result is a finite additive set function on the ring $\mathcal A$. In
order to use Theorem~\ref{T:masseind}, we therefore have to show
$\sigma$-additivity. For this, we will use Lemma~\ref{l:stetigcompact}
below, i.e.\ inner regularity of $P$ with respect to compact
sets. Before, we will show useful alternative conditions for
$\sigma$-additivity, which do not make use of any topological
structure of the underlying space.

\begin{lemma}[$\sigma$-additivity and continuity at $\emptyset$]\label{P:stetigmass}
  Let $\mathcal R$ be a ring and $m:\mathcal R\to \mathbb R_+$
  additive. Then, the following are equivalent:
  \begin{enumerate}
    \item $m$ is $\sigma$-additive;
    \item $m$ is $\sigma$-sub-additive;
    \item $m$ is continuous from below, i.e.\ for $A, A_1, A_2,... \in
      \mathcal R$ with $A_1 \subseteq A_2 \subseteq \cdots$, we have
      $m(A) = \lim_{n\to\infty} m(A_n)$.
    \item $m$ is continuous from above in $\emptyset$, i.e.\ for $A_1,
      A_2,...\in\mathcal R$ with $A_1 \supseteq A_2 \supseteq ...$ and
      $\bigcap_{n=1}^\infty A_n = \emptyset$, we have
      $\lim_{n\to\infty} m(A_n)=0$.
  \end{enumerate}
\end{lemma}

@@

  \noindent 1.$\Rightarrow$3.: For $A_1, A_2,...$ as in 3., we write
  (with $A_0 := \emptyset$)
  $$ m(A) = m\Big( \biguplus_{n=1}^\infty A_n \setminus
  \Big(\bigcup_{l=1}^{k-1} A_l\Big)\Big) = \lim_{n\to\infty} m\Big(
  \biguplus_{k=1}^n A_k \setminus \Big(\bigcup_{l=1}^{k-1}
  A_l\Big)\Big) = \lim_{n\to\infty} m(A_n).$$

@@

  \noindent 3.$\Rightarrow$1.: Let $A_1, A_2,... \in \mathcal R$ be
  disjoint and $A = \biguplus_{n=1}^\infty A_n$. Set $B_n :=
  \bigcup_{k=1}^n A_k$, such that $B_1, B_2,...$ satisfy 3. Hence,
  $$ m\Big( \biguplus_{n=1}^\infty A_n\Big) = m\Big(
  \bigcup_{n=1}^\infty B_n\Big) = \lim_{n\to\infty} m\Big(
  \bigcup_{k=1}^n B_n\Big) = \lim_{n\to\infty} m\Big(
  \biguplus_{k=1}^n A_n\Big) = \sum_{n=1}^\infty m(A_n).$$

@@

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  theorem countably_additive_of_todo (hC : SetRing C)
    (m : ∀ s : Set α, s ∈ C → ℝ≥0∞)
    (hm_ne_top : ∀ {s} (hs : s ∈ C), m s hs ≠ ∞)
    (hm_add : ∀ {s t : Set α} (hs : s ∈ C) (ht : t ∈ C),
      Disjoint s t →
      m (s ∪ t) (hC.union_mem hs ht) = m s hs + m t ht)
    (hm : ∀ ⦃s : ℕ → Set α⦄ (hs : ∀ n, s n ∈ C),
      Antitone s → (⋂ n, s n) = ∅ →
      Tendsto (fun n => m (s n) (hs n)) atTop (𝓝 0))
      ⦃f : ℕ → Set α⦄ (h : ∀ i, f i ∈ C) (hUf : (⋃ i, f i) ∈ C)
      (h_disj : Pairwise (Disjoint on f)) :
    m (⋃ i, f i) hUf = ∑' i, m (f i) (h i)
\end{minted}

@@

\begin{definition}[Inner regularity] \label{def:innerreg}
  Let $\alpha$ be some set, equipped with a topology, and $m$ be a
  set-function on some $\mathcal H \subseteq 2^\alpha$.
  \begin{enumerate}
  \item Let $p, q : 2^\alpha \to \{\text{true, false}\}$. Then, $m$ is
    called inner regular with respect to $p$ and $q$, if 
    $$ m(A) = \sup\{m(F) : p(F) = \text{true}, F \subseteq A\}$$ for
    all $A \in \mathcal H$ with $q(A) = \text{true}$.
  \item If $q(A) = \text{true}$ iff $A$ is measurable, we neglect the
    {\em and $q$}. If $p(A) = \text{true}$ iff $A$ is closed (ompact,
    closed and compact), we say that $m$ is inner regular with respect
    to closed (compact, compact and closed) sets.
  \end{enumerate}
\end{definition}

@@

For the next result, recall that for compact sets $C_1, C_2,...$ with
$\bigcap_{n=1}^\infty C_n = \emptyset$, there is some $N$ with
$\bigcap_{n=1}^N C_n = \emptyset$.

\begin{lemma}\label{l:stetigcompact}
  Let $\alpha$ be a topological space and $\mu$ be an additive set
  function on a ring $\mathcal R$ contained by the Borel
  $\sigma$-algebra, and which is inner regular with respect to compact
  sets. Then, $\mu$ is $\sigma$-additive.
\end{lemma}

\begin{proof}
  According to Lemma~\ref{P:stetigmass}, we need to show continuity of
  $\mu$ in $\emptyset$, so let $A_1 \supseteq A_2 \supseteq... \in
  \mathcal R$ satisfy $\bigcap_{n=1}^\infty = \emptyset$ and
  $\varepsilon>0$. Let $\delta_1, \delta_2,..>0$ with
  $\sum_{n=1}^\infty \delta_n < \varepsilon$. For each $n$, let $C_n
  \subseteq A_n \in \mathcal R$ be compact with $\mu(A_n) \leq
  \mu(C_n) + \delta_n$. We have that $\bigcap_{n=1}^\infty C_n
  \subseteq \bigcap_{n=1}^\infty A_n = \emptyset$, so there is $N \in
  \mathbb N$ with $\bigcap_{n=1}^N C_n = \emptyset$. So, for any
  $m>N$, we have that
  $$ \mu(A_m) = \mu\Big( \Big(\bigcap_{n=1}^m A_n\Big) \setminus
  \Big(\bigcap_{n=1}^m C_n\Big)\Big) \leq \sum_{n=1}^m \mu(A_n
  \setminus C_n) \leq \sum_{n=1}^m \delta_n < \varepsilon.$$ This
  concludes the proof.
\end{proof}

Here is the formalization using a \leanline{kol_content}:

@@

\sloppy Using this, we also have
\leanline{kolContent_countably_subadditive_of_innerRegular} by using
\leanline{countably_subadditive_of_countably_additive}.

Recall that a finite measure on the Borel $\sigma$-algebra is inner
regular with respect to closed sets, but also outer regular with
respect to open sets.

\begin{lemma}\label{l:pseude1}
  Let $r$ be an extended pseudo-metric on $\alpha$, and the topology
  on $\alpha$ be given by $r$. If $\mu$ is a finite measure on the
  Borel $\sigma$-algebra $\mathcal B$ on $\alpha$, it is inner regular
  with respect to closed sets. In fact, we have also outer regularity
  with respect to open sets, i.e.
  $$ \mu(A) = \inf \{ \mu(O) : A \subseteq O \text{ open}\}.$$
\end{lemma}

\begin{proof}
  It suffices to prove that
  $$ \mathcal A := \Big\{A \in \mathcal B : \mu(B) = \sup_{F\subseteq
    B \text{ closed}} \mu(F) = \sup_{B\subseteq O \text{ open}}
  \mu(O)\Big\} $$ is a $\sigma$-algebra containing all closed
  sets. So, we proceed in two steps.  \\ {\em Step 1: $\mathcal A$ is
    a $\sigma$-algebra:} Since $\mu$ is finite, we have that $\mu(A^c)
  = \mu(\alpha) - \mu(A)$ for all $A$. Using this, we already have
  that $\mathcal A$ is closed under complements. So, we are left with
  showing that $\mathcal A$ is closed under countable unions. For
  this, let $A_1, A_2,... \in \mathcal A$ and $A :=
  \bigcup_{n=1}^\infty A_n$. Fix $\varepsilon > 0$ and a sequence
  $\delta_1, \delta_2,... > 0$ with $\sum_{n=1}^\infty \delta_n <
  \varepsilon$. For each $n$, let $F_n \subseteq A_n \subseteq O_n$
  with $F_n$ closed, $O_n$ open and $\mu(O_n \setminus F_n) \leq
  \delta_n$. Then, $O := \bigcup_{n=1}^\infty O_n$ is open and
  $$ 0 \leq \mu(O \setminus A) = \mu(O) - \mu(A) \leq
  \sum_{n=1}^\infty \mu(O_n \setminus A) \leq \sum_{n=1}^\infty
  \mu(O_n \setminus A_n) < \varepsilon.$$ This shows outer regularity
  of $\mu$ with respect to open sets at $A$. It remains to show inner
  regularity with respect to closed sets. For this, note that
  $\bigcup_{n=1}^N F_n$ is closed for all $N$ and recall from
  Proposition~\ref{P:stetigmass} that $\mu$ is continuous from
  below. Therefore, setting $F := \bigcup_{n=1}^\infty F_n$,
  $$ 0 \leq \mu(A) - \lim_{N\to\infty} \mu\Big(\bigcup_{n=1}^N
  F_n\Big) = \mu(A \setminus F) \leq \sum_{n=1}^\infty \mu(A_n
  \setminus F) \leq \sum_{n=1}^\infty \mu(A_n \setminus F_n) <
  \varepsilon.$$ This shows inner regularity of $\mu$ with respect to
  closed sets at $A$.

  \noindent
      {\em Step 2: $\mathcal A$ contains all closed sets:} Let
  $\varepsilon_ n\downarrow 0$ and, for $\varepsilon>0$ and some $A
  \subseteq \alpha$, $A^\varepsilon := \{x : \exists y \in A,
  r(x,y<\varepsilon\}$. Then, for $A$ closed, we have that $A =
  \bigcap_{n=1}^\infty A^{\varepsilon_n}$ and $A^{\varepsilon_n}$ is
  open for all $n$. Clearly, $\mu(A) = \sup\{\mu(F): F\text{ closed},
  F \subseteq A\}$. By continuity of $\mu$, we find that
  $\mu(A^{\varepsilon_n}) \xrightarrow{n\to\infty} \mu(A)$, therefore
  $\mu(A) = \inf\{\mu(O): F\text{ open}, A \subseteq O\}$, and we have
  shown that $A \in \mathcal A$.
\end{proof}

For the formalization, we use a new result: \leanline{P.InnerRegular
  (fun s => IsCompact s ∧ IsClosed s) IsClosed}, which holds for every
second countable extended pseudo metric space, which is
\leanline{OpensMeasurable}. Then, from \leanline{mathlib}, we use
\leanline{Measure.InnerRegular.of_pseudoEMetricSpace} in order to show
that every open set is inner regular with respect to closed sets, and
\leanline{Measure.InnerRegular.measurableSet_of_open} to show the
above lemma.

@@

In the next lemma, we will need that a closed subset of a compact set
is compact. (See \leanline{isCompact_of_isClosed_subset}.)

@@

\begin{proof}
 Since $\alpha$ is separable, there is a subset $\{x_1, x_2,...\}
 \subseteq_c \alpha$ with closure $\alpha$. By separability, for each
 $n\in \mathbb N$, $\alpha = \bigcup_{k=1}^\infty
 B_{\varepsilon_n}(x_k)$. Since $\mu$ is continuous from above in
 $\emptyset$ (see Lemma~\ref{P:stetigmass})
  \[
  0 = \mu\Big(\Big(\bigcup_{k=1}^\infty B_{\varepsilon_n}(x_k)\Big)^c
  \Big) = \lim_{N\to\infty} \mu\Big( \Big(\bigcup_{k=1}^N
  B_{\varepsilon_n}(x_k)\Big)^c\Big).
  \]
  Let $\delta_n\downarrow 0$ be summable with $\sum_{n=1}^\infty
  \delta_n = 1$ (e.g.\ $\delta_n = 2^{-n}$). Then, there is some
  $N_n\in\mathbb N$ with $\mu\Big( E\setminus \bigcup_{k=1}^{N_n}
  B_{\varepsilon_n}(x_k)\Big) < \varepsilon \delta_n$. Now take
  \[
    A := \bigcap_{n=1}^\infty \bigcup_{k=1}^{N_n}
    B_{\varepsilon_n}(x_k),
  \]
  which by construction is totally bounded (for $r>0$ take $n$ such
  that $\varepsilon_n < r$, such that $B_r(x_1),...,B_r(x_{N_n})$
  cover $A$), and therefore has a compact closure as noted before the
  lemma. In addition, by $\sigma$-sub-additivity of $\mu$,
  $$ \mu((\overline A)^c) \leq \mu(A^c) = \mu\Big(
  \bigcup_{n=1}^\infty \Big(\Big(\bigcup_{k=1}^{N_n}
  B_{\varepsilon_n}(x_k)\Big)^c\Big)\Big) \leq \sum_{n=1}^\infty
  \mu\Big( \Big(\bigcup_{k=1}^{N_n}
  B_{\varepsilon_n}(x_k)\Big)^c\Big)< \varepsilon,$$ and we are done.
\end{proof}

Let us combine the last two results in one combined formalized result:

@@

\noindent
With the above results, the strategy of the proof of Theorem~\ref{T1}
is as follows: We have to show that under the given conditions,
condition~1 of Lemma~\ref{l:tight} holds. By the equivalent~2., we
find that $P$ is inner regular with respect to closed compact sets. In
particular, the conditions of Lemma~\ref{l:stetigcompact} are
satisfied, showing that $P$ is $\sigma$-additive. Then, we proceed as
described right before Proposition~\ref{P:stetigmass}. So, it remains
to show condition~1 of Lemma~\ref{l:tight} for complete and separable
extended pseudo-metric spaces. Note that -- as a special case of
\leanline{isCompact_iff_totallyBounded_isComplete} from {\tt
  mathlib}, any subset of such a space is compact iff it is complete
and totally bounded. Since closed subsets of a complete space are
complete, the closure of a totally bounded set is (still totally
bounded, hence) compact.

@@

\subsection{Proof of Kolmogorov's extension theorem}
Now we can put everything together: According to
Lemma~\ref{L:relcoPol}, condition~1 of Lemma~\ref{l:tight} is
satisfied for the set function $P$ on the ring $\mathcal A$ from
Theorem~\ref{T1}. The same lemma shows that the conditions for
Lemma~\ref{l:stetigcompact} are satisfied, so $P$ is $\sigma$-additive
on $\mathcal A$. As a consequence, we can use the Carathéodory
extension from Theorem~\ref{T:masseind}. This shows all other
assertions. The formalization of this proof resembles these
arguments. We leave out all instances in the reformulation of the
result and its proof (see below Theorem~\ref{T1} for a full
formulation):

@@

\section{Concluding thoughts}
{\em Using semi-rings in measure theory:} Frequently, the construction
of measures on some $\sigma$-algebra $\mathcal F$ (on some space
$\alpha$) uses outer measures, which can be defined on $2^\alpha$ (the
set of all subsets of $\alpha$). We provide a general framework using
Carathéodory's extension theorem, which states an extension of a
set-function on a semi-ring by an outer measure, not previously
implemented in mathlib. This could also be used to re-define
Stieltjes measures (in particular Lebesgue measure) on $\mathbb R$, as
well as product measures (on finite products). More precisely,
e.g.\ \leanline{StieltjesFunction.outer_Ioc} and
\leanline{MeasureTheory.constructions.pi.pi_pi_aux} are specific
examples of the more general statement, that the outer measure extends
the set function a the semi-ring. (In the first case, it would be the
semi-ring of half-open intervals, in the second case, it would be the
semi-ring of cylinder sets on a product space.)

@@

xxx design decisions?

>>>>

\begin{theorem}[Kolmogorov extension]\label{T1}
  For all $t\in\iota$, let $\alpha_t$ be a separable, complete
  pseudo-extended-metric space and $\mathcal F_t$ the Borel
  $\sigma$-algebra generated by its topology. Let $(P_J)_{J\subseteq_f
    \iota}$ be a projective family of finite measures and $P$ be
  defined on $\mathcal A := \bigcup_{J \subseteq_f \iota} \mathcal
  F_J$ given by $P(A) = P_J(\pi_J A)$ for $A\in\mathcal F_J$. Then,
  there is a unique extension of $P$ to $\sigma(\mathcal A)$.
\end{theorem}

@@

\noindent
In the proof of Kolmogorov's extension theorem, we use a closely
related definition of \leanline{Measure.ofAddContent}, with variables
\leanline{[mα : MeasurableSpace α] (hC : SetSemiring C) (hC_gen :
  MeasurableSpace.generateFrom C = mα)}, i.e.\ the semi-ring
\leanline{C} generates the $\sigma$-algebra, and which uses an
\leanline{m : AddContent} as well as \leanline{m_sigma_subadd} as
above.

\subsection{$\sigma$-additivity of set functions}
For the proof of Kolmogorov's extension Theorem, note that $P$ as
given in Theorem~\ref{T1} is a finite additive set function on the
ring $\mathcal A$. In order to use Theorem~\ref{T:masseind}, we
therefore have to show $\sigma$-additivity. For this, we will use
Lemma~\ref{l:stetigcompact} below, i.e.\ inner regularity of $P$ with
respect to compact sets. Before, we will show useful alternative
conditions for $\sigma$-additivity, which do not make use of any
topological structure of the underlying space.

\begin{lemma}[$\sigma$-additivity and continuity at $\emptyset$]\label{P:stetigmass}
  Let $\mathcal R$ be a ring and $m:\mathcal R\to \mathbb R_+$
  additive. Then, the following are equivalent:
  \begin{enumerate}
    \item $m$ is $\sigma$-additive;
    \item $m$ is $\sigma$-sub-additive;
    \item $m$ is continuous from below, i.e.\ for $A, A_1, A_2,... \in
      \mathcal R$ with $A_1 \subseteq A_2 \subseteq \cdots$ and $A =
      \bigcup_{n=1}^\infty A_n$, we have $m(A) = \lim_{n\to\infty}
      m(A_n)$.
    \item $m$ is continuous from above in $\emptyset$, i.e.\ for $A_1,
      A_2,...\in\mathcal R$ with $A_1 \supseteq A_2 \supseteq ...$ and
      $\bigcap_{n=1}^\infty A_n = \emptyset$, we have
      $\lim_{n\to\infty} m(A_n)=0$.
  \end{enumerate}
\end{lemma}

@@

  \noindent 1.$\Rightarrow$3.: For $A_1, A_2,...$ as in 3., we write
  (with $A_0 := \emptyset$)
  \begin{align*}
     m(A) & % = m\Big( \biguplus_{n=1}^\infty \Big(A_n \setminus
     %\Big(\bigcup_{l=1}^{k-1} A_l\Big)\Big)\Big)
     = \lim_{n\to\infty}
     m\Big( \biguplus_{k=1}^n A_k \setminus \Big(\Big(\bigcup_{l=1}^{k-1}
     A_l\Big)\Big)\Big) = \lim_{n\to\infty} m(A_n).
  \end{align*}

@@

  \noindent 3.$\Rightarrow$1.: Let $A_1, A_2,... \in \mathcal R$ be
  disjoint and $A = \biguplus_{n=1}^\infty A_n$. Set $B_n :=
  \bigcup_{k=1}^n A_k$, such that $B_1, B_2,...$ satisfy 3. Hence,
  \begin{align*}
    m\Big( \biguplus_{n=1}^\infty A_n\Big) & = m\Big(
    \bigcup_{n=1}^\infty B_n\Big) = \lim_{n\to\infty} m\Big(
    \bigcup_{k=1}^n B_n\Big) \\ & = \lim_{n\to\infty} m\Big(
    \biguplus_{k=1}^n A_n\Big) = \sum_{n=1}^\infty m(A_n).
  \end{align*}

@@

xxx Rename the todo!

\begin{minted}[mathescape, numbersep=5pt, framesep=5mm, bgcolor=mygray]{Lean}
  theorem countably_additive_of_todo (hC : SetRing C)
    (m : ∀ s : Set α, s ∈ C → ℝ≥0∞)
    (hm_ne_top : ∀ {s} (hs : s ∈ C), m s hs ≠ ∞)
    (hm_add : ∀ {s t : Set α} (hs : s ∈ C) (ht : t ∈ C),
      Disjoint s t →
      m (s ∪ t) (hC.union_mem hs ht) = m s hs + m t ht)
    (hm : ∀ ⦃s : ℕ → Set α⦄ (hs : ∀ n, s n ∈ C),
      Antitone s → (⋂ n, s n) = ∅ →
      Tendsto (fun n => m (s n) (hs n)) atTop (nhds 0))
      ⦃f : ℕ → Set α⦄ (h : ∀ i, f i ∈ C) (hUf : (⋃ i, f i) ∈ C)
      (h_disj : Pairwise (Disjoint on f)) :
    m (⋃ i, f i) hUf = ∑' i, m (f i) (h i)
\end{minted}

@@

\begin{definition}[Inner regularity] \label{def:innerreg}
  Let $\alpha$ be some set, equipped with a topology, and $m$ be a
  set-function on some $\mathcal H \subseteq 2^\alpha$.
  \begin{enumerate}
  \item Let $p, q : 2^\alpha \to \{\text{true, false}\}$. Then, $m$ is
    called inner regular with respect to $p$ and $q$, if 
    $$ m(A) = \sup\{m(F) : p(F) = \text{true}, F \subseteq A\}$$ for
    all $A \in \mathcal H$ with $q(A) = \text{true}$.
  \item If $q(A) = \text{true}$ iff $A$ is measurable, we neglect the
    {\em and $q$}. If $p(A) = \text{true}$ iff $A$ is closed (compact,
    closed and compact), we say that $m$ is inner regular with respect
    to closed (compact, compact and closed) sets.
  \end{enumerate}
\end{definition}

@@

For the next result, recall that for compact sets $C_1, C_2,...$ with
$\bigcap_{n=1}^\infty C_n = \emptyset$, there is some $N$ with
$\bigcap_{n=1}^N C_n = \emptyset$. More generally, compact sets form a
compact system, which is defined as follows:

\begin{definition}
  Let $\mathcal C \subseteq 2^\alpha$. If, for all $C_1, C_2,...$ with
  $\bigcap_{n=1}^\infty C_n = \emptyset$, there is some $N$ with
  $\bigcap_{n=1}^N C_n = \emptyset$, we call $\mathcal C$ a {\em
    compact system}.
\end{definition}

\begin{lemma}\label{l:stetigcompact}
  Let $\alpha$ be a topological space and $\mu$ be an additive set
  function on a ring $\mathcal R$ contained by the Borel
  $\sigma$-algebra, and which is inner regular with respect to a
  compact system. Then, $\mu$ is $\sigma$-additive.
\end{lemma}

\begin{proof}
  According to Lemma~\ref{P:stetigmass}, we need to show continuity of
  $\mu$ in $\emptyset$, so let $A_1 \supseteq A_2 \supseteq... \in
  \mathcal R$ satisfy $\bigcap_{n=1}^\infty A_n = \emptyset$ and
  $\varepsilon>0$. Let $\delta_1, \delta_2,... >0$ with
  $\sum_{n=1}^\infty \delta_n < \varepsilon$. For each $n$, let
  $\mathcal C \ni C_n \subseteq A_n \in \mathcal R$ with $\mu(A_n)
  \leq \mu(C_n) + \delta_n$. We have that $\bigcap_{n=1}^\infty C_n
  \subseteq \bigcap_{n=1}^\infty A_n = \emptyset$, so there is $N \in
  \mathbb N$ with $\bigcap_{n=1}^N C_n = \emptyset$ since $\mathcal C$
  is a compact system. So, for any $m>N$, we have that
  \begin{align*}
     \mu(A_m) & = \mu\Big( \Big(\bigcap_{n=1}^m A_n\Big) \setminus
     \Big(\bigcap_{n=1}^m C_n\Big)\Big) \leq \sum_{n=1}^m \mu(A_n
     \setminus C_n) \\ & \leq \sum_{n=1}^m \delta_n < \varepsilon.
  \end{align*}
  This concludes the proof.
\end{proof}

xxx Work on this, of time permits!

In our formalization, we are less general and use a direct argument
for a \leanline{kol_content}. More precisely, we show that a
\leanline{kolContent}, which is based on a projective family which is
inner regular with respect to compact and closed sets is
$\sigma$-additive.

@@

For using the more general version, we would need to show that
$\{\pi_J^{-1} C: C \in \prod_{j\in J} \alpha_j \text{ compact}\}$ as
well as $\{\pi_J^{-1} C: C \in \prod_{j\in J} \alpha_j \text{ compact
  and closed}\}$ are compact systems. (Note that compact sets are
closed in Hausdorff spaces, but we do not have this property since we
are working with pseudo-metric spaces.) In addition, the assumption
\leanline{hP_inner} from above translates directly to inner regularity
of the \leanline{kolContent}, which is defined through the projective
family \leanline{P}. Then, Lemma~\ref{l:stetigcompact} gives the same
result as the formalization. Moreover, since the Carathéodory
extension theorem requires $\sigma$-subadditivity (rather than
$\sigma$-additivity), we can use
\leanline{countably_subadditive_of_countably_additive} in order to
show \leanline{kolContent_countably_subadditive_of_innerRegular},
which shows that the assumptions of the last result also imply
$\sigma$-subadditivity.

What remains to be done is to show conditions under which the
projective family is inner regular with respect to compact (and
closed) sets. For this, we need some assumption on the underlying
spaces. The following lemma provides the generalizations to extended
pseudo-metric spaces:

The next result is already implemented in \leanline{mathlib}: Use
\leanline{Measure.InnerRegular.of_pseudoEMetricSpace} in order to show
that every open set is inner regular with respect to closed sets, and
\leanline{Measure.InnerRegular.measurableSet_of_open} to show that
every measurable set is inner regular with respect to closed sets.
The claimed outer reularity is part of
\leanline{MeasureTheory.Measure.InnerRegular.weaklyRegular_of_finite}.

\begin{lemma}\label{l:pseude1}
  Let $r$ be an extended pseudo-metric on $\alpha$, and the topology
  on $\alpha$ be given by $r$. If $\mu$ is a finite measure on the
  Borel $\sigma$-algebra $\mathcal B$ on $\alpha$, it is inner regular
  with respect to closed sets. In fact, we have also outer regularity
  with respect to open sets, i.e.\ for all $A\in \mathcal B$,
  $$ \mu(A) = \inf \{ \mu(O) : A \subseteq O \text{ open}\}.$$
\end{lemma}

\begin{proof}
  It suffices to prove that
  $$ \mathcal A := \Big\{A \in \mathcal B : \mu(B) = \sup_{F\subseteq
    B \text{ closed}} \mu(F) = \sup_{B\subseteq O \text{ open}}
  \mu(O)\Big\} $$ is a $\sigma$-algebra containing all closed
  sets. So, we proceed in two steps.  \\ {\em Step 1: $\mathcal A$ is
    a $\sigma$-algebra:} Since $\mu$ is finite, we have that $\mu(A^c)
  = \mu(\alpha) - \mu(A)$ for all $A$. Using this, we already have
  that $\mathcal A$ is closed under complements. So, we are left with
  showing that $\mathcal A$ is closed under countable unions. For
  this, let $A_1, A_2,... \in \mathcal A$ and $A :=
  \bigcup_{n=1}^\infty A_n$. Fix $\varepsilon > 0$ and a sequence
  $\delta_1, \delta_2,... > 0$ with $\sum_{n=1}^\infty \delta_n <
  \varepsilon$. For each $n$, let $F_n \subseteq A_n \subseteq O_n$
  with $F_n$ closed, $O_n$ open and $\mu(O_n \setminus F_n) \leq
  \delta_n$. Then, $O := \bigcup_{n=1}^\infty O_n$ is open and
  $$ 0 \leq \mu(O) - \mu(A) \leq \sum_{n=1}^\infty \mu(O_n \setminus
  A) \leq \sum_{n=1}^\infty \mu(O_n \setminus A_n) < \varepsilon.$$
  This shows outer regularity of $\mu$ with respect to open sets at
  $A$. It remains to show inner regularity with respect to closed
  sets. For this, note that $\bigcup_{n=1}^N F_n$ is closed for all
  $N$ and recall from Proposition~\ref{P:stetigmass} that $\mu$ is
  continuous from below. Therefore, setting $F := \bigcup_{n=1}^\infty
  F_n$,
  \begin{align*}
    0 & \leq \mu(A) - \lim_{N\to\infty} \mu\Big(\bigcup_{n=1}^N
    F_n\Big) = \mu(A \setminus F) \leq \sum_{n=1}^\infty \mu(A_n
    \setminus F) \\ & \leq \sum_{n=1}^\infty \mu(A_n \setminus F_n) <
    \varepsilon.
  \end{align*}
  This shows inner regularity of $\mu$ with respect to closed sets at
  $A$.

  \noindent
  {\em Step 2: $\mathcal A$ contains all closed sets:} Let
  $\varepsilon_ n\downarrow 0$ and, for $\varepsilon>0$ and some $A
  \subseteq \alpha$, $A^\varepsilon := \{x : \exists y \in A,
  r(x,y)<\varepsilon\}$. Then, for $A$ closed, we have that $A =
  \bigcap_{n=1}^\infty A^{\varepsilon_n}$ and $A^{\varepsilon_n}$ is
  open for all $n$. Clearly, $\mu(A) = \sup\{\mu(F): F\text{ closed},
  F \subseteq A\}$, since $A$ is closed. By continuity of $\mu$, we
  find that $\mu(A^{\varepsilon_n}) \xrightarrow{n\to\infty} \mu(A)$,
  therefore $\mu(A) = \inf\{\mu(O): F\text{ open}, A \subseteq O\}$,
  and we have shown that $A \in \mathcal A$.
\end{proof}

@@

In the next lemma, we will need that a closed subset of a compact set
is compact. (See \leanline{isCompact_of_isClosed_subset}.) We describe
the formalization of the next lemma only in conjunction with
Lemma~\ref{L:relcoPol}.

@@

\begin{proof}
 Since $\alpha$ is separable, there is a subset $\{x_1, x_2,...\}
 \subseteq_c \alpha$ with closure $\alpha$. By separability, for each
 $n\in \mathbb N$, $\alpha = \bigcup_{k=1}^\infty
 B_{\varepsilon_n}(x_k)$. Since $\mu$ is continuous from above in
 $\emptyset$ (see Lemma~\ref{P:stetigmass})
  \[
  0 = \mu\Big(\Big(\bigcup_{k=1}^\infty B_{\varepsilon_n}(x_k)\Big)^c
  \Big) = \lim_{N\to\infty} \mu\Big( \Big(\bigcup_{k=1}^N
  B_{\varepsilon_n}(x_k)\Big)^c\Big).
  \]
  Let $\delta_n\downarrow 0$ be summable with $\sum_{n=1}^\infty
  \delta_n = 1$ (e.g.\ $\delta_n = 2^{-n}$). Then, there is some
  $N_n\in\mathbb N$ with $\mu\Big( E\setminus \bigcup_{k=1}^{N_n}
  B_{\varepsilon_n}(x_k)\Big) < \varepsilon \delta_n$. Now take
  \[
    A := \bigcap_{n=1}^\infty \bigcup_{k=1}^{N_n}
    B_{\varepsilon_n}(x_k),
  \]
  which by construction is totally bounded (for $r>0$ take $n$ such
  that $\varepsilon_n < r$, such that $B_r(x_1),...,B_r(x_{N_n})$
  cover $A$), and therefore has a compact closure as noted before the
  lemma. In addition, by $\sigma$-sub-additivity of $\mu$,
  \begin{align*}
     \mu((\overline A)^c) & \leq \mu(A^c) = \mu\Big(
     \bigcup_{n=1}^\infty \Big(\Big(\bigcup_{k=1}^{N_n}
     B_{\varepsilon_n}(x_k)\Big)^c\Big)\Big) \\ & \leq
     \sum_{n=1}^\infty \mu\Big( \Big(\bigcup_{k=1}^{N_n}
     B_{\varepsilon_n}(x_k)\Big)^c\Big)< \varepsilon,
  \end{align*}
 and we are done.
\end{proof}

Let us combine the last two results in one combined formalized result.
This the gives the desired result that a measure on a second
countable, complete extended pseudo-metric space is inner regular with
respect to closed compact sets.

@@

@@

\subsection{Proof of Kolmogorov's extension theorem}
With the above results, the strategy of the proof of Theorem~\ref{T1}
is as follows: We have to show that under the given conditions,
condition~1 of Lemma~\ref{l:tight} holds. By the equivalent~2., we
find that $P$ is inner regular with respect to closed compact sets. In
particular, the conditions of Lemma~\ref{l:stetigcompact} are
satisfied, showing that $P$ is $\sigma$-additive. Then, we proceed as
described right before Proposition~\ref{P:stetigmass}. So, it remains
to show condition~1 of Lemma~\ref{l:tight} for complete and separable
extended pseudo-metric spaces. Note that -- as a special case of
\leanline{isCompact_iff_totallyBounded_isComplete} from {\tt
  mathlib}, any subset of such a space is compact iff it is complete
and totally bounded. Since closed subsets of a complete space are
complete, the closure of a totally bounded set is (still totally
bounded, hence) compact.

Now we can put everything together: According to
Lemma~\ref{L:relcoPol}, condition~1 of Lemma~\ref{l:tight} is
satisfied for the set function $P$ on the ring $\mathcal A$ from
Theorem~\ref{T1}. The same lemma shows that the conditions for
Lemma~\ref{l:stetigcompact} are satisfied, so $P$ is $\sigma$-additive
on $\mathcal A$. As a consequence, we can use the Carathéodory
extension from Theorem~\ref{T:masseind}. This shows all other
assertions. The formalization of this proof resembles these
arguments. We leave out all instances in the reformulation of the
result and its proof (see below Theorem~\ref{T1} for a full
formulation):

@@

\section{Concluding thoughts}
We will highlight some paths we imagine can be taken, leading to
further extensions of \leanline{mathlib}, which build on our
contribution.

{\em Using semi-rings in measure theory:} Frequently, the construction
of measures on some $\sigma$-algebra $\mathcal F$ (on some space
$\alpha$) uses outer measures, which can be defined on $2^\alpha$ (the
set of all subsets of $\alpha$). We provide a general framework using
Carathéodory's extension theorem, which states an extension of a
set-function on a semi-ring by an outer measure, not previously
implemented in mathlib. This could also be used to re-define Stieltjes
measures (in particular Lebesgue measure) on $\mathbb R$, as well as
product measures (on finite products). More precisely,
e.g.\ \leanline{StieltjesFunction.outer_Ioc} and
\leanline{MeasureTheory.constructions.pi.pi_pi_aux} are specific
examples of the more general statement, that the outer measure extends
the set function a the semi-ring. (In the first case, it would be the
semi-ring of half-open intervals, in the second case, it would be the
semi-ring of cylinder sets on a product space.)

@@

~~
